Actuator line modeling of vertical-axis turbines

### 2016-05-12 paper.tex
Discuss wake transport ratios; resolves #20
Update corresponding email; resolves #22
Note that this is a preprint; resolves #23
Remove 'hybridization' from abstract

\articletype{RESEARCH ARTICLE}

@@

\corraddr{P.~Bachant, Chase Ocean Engineering Laboratory, University of New
Hampshire, 24 Colovos Rd., Durham, NH, 03824, USA. \\
E-mail: pxL3@unh.edu}

\begin{abstract}
    To bridge the gap between high and low fidelity numerical modeling tools for
    vertical-axis (or cross-flow) turbines (VATs or CFTs), an actuator line
    model (ALM) was developed and validated for both a high and a medium
    solidity vertical-axis turbine at rotor diameter Reynolds numbers $Re_D \sim
    10^6$. The ALM is a hybridization of classical blade element theory with
    Navier--Stokes based flow models, and in this study both $k$--$\epsilon$
    Reynolds-averaged Navier--Stokes (RANS) and Smagorinsky large eddy
    simulation (LES) turbulence models were tested. The RANS models were able to
    be run on coarse grids while still providing good convergence behavior in
    terms of the mean power coefficient, and also approximately four orders of
    magnitude reduction in computational expense compared with 3-D
    blade-resolved RANS simulations. Submodels for dynamic stall, end effects,
    added mass, and flow curvature were implemented, resulting in reasonable
    performance predictions for the high solidity rotor, more discrepancies for
    the medium solidity rotor, and overprediction for both cases at high tip
    speed ratio. The wake results showed that the ALM was able to capture some
    of the important flow features that contribute to VAT's relatively fast wake
    recovery---a large improvement over the conventional actuator disk model.
    The mean flow field was better realized with the LES, which still
    represented a computational savings of two orders of magnitude compared with
    3-D blade-resolved RANS, though vortex breakdown and subsequent turbulence
    generation appeared to be underpredicted, which necessitates further
    investigation of optimal subgrid scale modeling.
\end{abstract}

@@

Finally, a similar mean streamwise momentum transport analysis was undertaken by
computing weighted sums of each term across the entire domain in the $y$--$z$
directions. Similar results as for the UNH-RVAT were obtained, i.e.,
cross-stream advection was predicted to be positive where it should have been
negative, vertical advection was predicted reasonably well, and turbulent
transport due to the eddy viscosity was also relatively large.

>>>>

\articletype{RESEARCH ARTICLE PREPRINT (submitted May 2016)}

@@

\corraddr{P.~Bachant, Chase Ocean Engineering Laboratory, University of New
Hampshire, 24 Colovos Rd., Durham, NH, 03824, USA. \\
E-mail: pwbachant@gmail.com}

\begin{abstract}
    To bridge the gap between high and low fidelity numerical modeling tools for
    vertical-axis (or cross-flow) turbines (VATs or CFTs), an actuator line
    model (ALM) was developed and validated for both a high and a medium
    solidity vertical-axis turbine at rotor diameter Reynolds numbers $Re_D \sim
    10^6$. The ALM is a combination of classical blade element theory and
    Navier--Stokes based flow models, and in this study both $k$--$\epsilon$
    Reynolds-averaged Navier--Stokes (RANS) and Smagorinsky large eddy
    simulation (LES) turbulence models were tested. The RANS models were able to
    be run on coarse grids while still providing good convergence behavior in
    terms of the mean power coefficient, and also approximately four orders of
    magnitude reduction in computational expense compared with 3-D
    blade-resolved RANS simulations. Submodels for dynamic stall, end effects,
    added mass, and flow curvature were implemented, resulting in reasonable
    performance predictions for the high solidity rotor, more discrepancies for
    the medium solidity rotor, and overprediction for both cases at high tip
    speed ratio. The wake results showed that the ALM was able to capture some
    of the important flow features that contribute to VAT's relatively fast wake
    recovery---a large improvement over the conventional actuator disk model.
    The mean flow field was better realized with the LES, which still
    represented a computational savings of two orders of magnitude compared with
    3-D blade-resolved RANS, though vortex breakdown and subsequent turbulence
    generation appeared to be underpredicted, which necessitates further
    investigation of optimal subgrid scale modeling.
\end{abstract}

@@

Finally, a similar mean streamwise momentum transport analysis was undertaken by
computing weighted sums of each term across the entire domain in the $y$--$z$
directions. Similar results as for the UNH-RVAT were obtained, i.e.,
cross-stream advection was predicted to be positive where it should have been
negative, vertical advection was predicted reasonably well, and turbulent
transport due to the eddy viscosity was also relatively large. The ratio of wake
transport compared with the UNH-RVAT RANS case (approximately 60\% to 70\% lower
for the RM2) matches well with that computed from the experiments, which shows
the ALM may successfully predict larger optimal array spacing for the RM2 versus
UNH-RVAT.
